```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(GenomicRanges)
ROOT_DIR <- "/mnt/isilon/sfgi/dudekm/"
```

In this notebook, I will create a master table of fpQTLs which combines multiple different types of annotations:
- distance to nearest TSS, name of nearest gene
- list of associated traits (GWAS)
- list of associated genes (liver eQTL)
- list of overlapping ChIP-peaks (liver, HepG2, hepatocyte)
- list of overlapping concordant motifs
- liver ADASTRA associations

```{r}
FP_METHOD <- "PRINT_beta_logit"
MOTIF_THRESHOLD <- "p=5e-4"
```


# Load data
Load fpQTLs
```{r}
fpQTLs <- read.delim(paste0("../regression/FP_methods/", FP_METHOD, 
                            "/regression_results/fpQTLs_covariates_FDR5.txt"))

fpQTL.range <- GRanges(seqnames = fpQTLs$variant_chrom, 
                         ranges = IRanges(start = fpQTLs$variant_pos, 
                                          end = fpQTLs$variant_pos))
```

Load TSS ranges from RefSeq Curated
```{r}
refSeqCurated <- read.table(paste0(ROOT_DIR, "raw_data/genome/RefSeq/ncbiRefSeqCurated.txt"))

colnames(refSeqCurated) <- c("bin", "name", "chrom", "strand", "txStart", "txEnd", "cdsStart", "cdsEnd", "exonCount", "exonStarts", "exonEnds", "score", "name2", "cdsStartStat", "cdsEndStat", "exonFrames")

refSeqCurated %>%
  mutate(
    tx = ifelse(strand == "+", txStart, txEnd)
  ) ->
  refSeqCurated

tss.range <- GRanges(seqnames = refSeqCurated$chrom, 
                     ranges = IRanges(start = refSeqCurated$tx, 
                                      end = refSeqCurated$tx,
                                      names = refSeqCurated$name2))
```

Load GWAS and eQTL
```{r}
# proxy_union_GWsig <- readRDS(paste0(ROOT_DIR, "raw_data/GWAS/sentinel_variants/snipa_LDproxies_r2=0.8_1000G_Phase3v5/proxy_union_GWsignificant/proxy_union_GWsignificant.Rds"))
# 
# # get list of rsIDs for each trait
# significant_rsIDs <- list()
# for (trait in names(proxy_union_GWsig)) {
#   proxy_union_GWsig[[trait]] %>%
#     separate_rows(rsID_alias, sep = ",") ->
#     df
#   
#   significant_rsIDs[[trait]] <- c(proxy_union_GWsig[[trait]][["rsID_alias"]], 
#                                   proxy_union_GWsig[[trait]][["rsID"]]) %>% na.omit()
# }

R2 <- "0.8"

GWAS_DIR <- paste0(ROOT_DIR, "raw_data/GWAS/sentinel_variants/snipa_LDproxies_r2=", R2, "_1000G_Phase3v5/")
merged_proxies <- read.delim(paste0(GWAS_DIR, "merged/merged_proxies.txt"))

merged_proxies <- merged_proxies %>% dplyr::select(RSID, TRAIT)

# Load dbSNP151 rsID key - fixes some nonstandard rsIDs from Gencove
dbSNP151_rsKey <- readRDS(paste0(ROOT_DIR, "footprinting/fpQTLs/misc/variant_ids/gencove_vcf_to_dbSNP155_key.Rds"))


liver.eQTLs <- readRDS(paste0(ROOT_DIR, "raw_data/GTEx/Liver.v8.signif_variants.hg38.names.Rds"))

```

Load ChIP peaks
```{r}
chip_peaks <- read.delim(paste0(ROOT_DIR, "raw_data/ChIP-seq/tf_clusters_liver_ENCODE/new_hg38/hg38_tf_clusters_liver_labelled.txt"))

chip_peaks_range <- GRanges(seqnames = chip_peaks$chrom, 
                                   ranges = IRanges(start = chip_peaks$chromStart, 
                                                    end = chip_peaks$chromEnd))
```

Load motif concordant fpQTLs
```{r}
concordant_fpQTLs <- read.delim(paste0("../motif_analysis/concordant_fpQTLs/", MOTIF_THRESHOLD, 
                     "/", FP_METHOD, "_concordant_fpQTLs.txt"))

concordant_fpQTLs_sum <- concordant_fpQTLs %>%
  group_by(variant_id) %>%
  summarise(
    concordant_motif_ids = paste(motif_id, collapse = ","),
    concordant_motif_delta_scores = paste(delta_score, collapse = ",")
  )
```

Load ADASTRA
```{r}
ADASTRA <- read.delim(paste0(ROOT_DIR, "raw_data/ChIP-seq/ADASTRA/AD_ASTRA_v5.1.3_liver_HepG2_hepatocytes.tsv"))

ADASTRA <- ADASTRA %>%
  dplyr::select(rsID, TF.ASBs) %>%
  dplyr::rename(
    variant_id = rsID,
    ADASTRA_TFs = TF.ASBs
  )
```


# TSS annotation
```{r}
result <- distanceToNearest(fpQTL.range, tss.range)
nearest_index <- subjectHits(result)
fpQTLs$nearest_tss <- refSeqCurated$name2[nearest_index]
fpQTLs$distance_to_nearest_tss <- mcols(result)$distance
fpQTLs$signed_tss_distance <- ifelse(
  refSeqCurated$strand[nearest_index] == "+",
  fpQTLs$variant_pos - refSeqCurated$tx[nearest_index],
  refSeqCurated$tx[nearest_index] - fpQTLs$variant_pos
) 
fpQTLs$signed_tss_distance <- fpQTLs$signed_tss_distance - sign(fpQTLs$signed_tss_distance)
```

# GWAS annotation
```{r}
get_traits <- function(rsID1, rsID2) {
  traits <- merged_proxies %>% dplyr::filter(RSID == rsID1 | RSID == rsID2) %>% pull(TRAIT)
   if (length(traits) == 0) NA else traits[1]
}

fpQTLs <- fpQTLs %>%
  # Get dbSNP rsID, which might differ from original rsID
  left_join(dbSNP151_rsKey, by = join_by(variant_id == original_rsID)) %>% 
  separate(variant_id, into = c("variant_id", NA), sep = "_", fill = "right") %>% # "rsXXXXXX_2" --> "rsXXXXXX"
  mutate(
    rsID.dbSNP = ifelse(is.na(rsID.dbSNP), variant_id, rsID.dbSNP)
  ) %>%
  rowwise() %>%
  mutate(associated_traits = get_traits(variant_id, rsID.dbSNP)) %>%
  ungroup()
```


# eQTL annotation
```{r}
fpQTLs <- fpQTLs %>%
  left_join(liver.eQTLs, by = join_by(variant_chrom, variant_pos))
```

# ChIP-seq annotation
```{r}
overlaps <- GenomicRanges::findOverlaps(fpQTL.range, chip_peaks_range, select = "all")
  
data.frame(overlaps) %>%
  mutate(tf_name = chip_peaks$name[subjectHits]) %>%
  group_by(queryHits) %>%
  summarize(tf_names = paste(unique(tf_name), collapse = ",")) ->
  overlaps

fpQTLs$liver_TF_ChIP <- NA
fpQTLs$liver_TF_ChIP[overlaps$queryHits] <- overlaps$tf_names
```

# Concordant motif annotation
```{r}
fpQTLs <- fpQTLs %>%
  left_join(concordant_fpQTLs_sum, by = join_by(variant_id))
```

# ADASTRA annotation
```{r}
fpQTLs <- fpQTLs %>% 
  left_join(ADASTRA, by = join_by(variant_id))
```


```{r}
fpQTLs %>%
  write.table(paste0("fpQTLs_annotated/", FP_METHOD, "_annotated_fpQTLs.txt"),
              quote = FALSE, row.names = FALSE, sep = "\t")
```













